---
title: "A tutorial to run diffdriver"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```

# Environment requirement

Need 16G memory to run diffdriver. The place that need the most memory is for parameter estimation for background mutation rate.

# Prepare input

  * Gene file
```{r}
genef = system.file("extdata", "example_gene.txt", package = "diffdriver")
head(read.table(genef, header = F))
```

  * Mutation file
```{r}
mutf = system.file("extdata/", "example_mutations.txt", package = "diffdriver")
head(read.table(mutf, header = T)) 
```

  * Phenotype file
  
Note: Nsyn is only required for when running diffdriver with mode "regular". TODO: update regular BMR so that this is not used.
```{r}
phenof = system.file("extdata/", "example_phenotypes.txt", package = "diffdriver")
head(read.table(phenof, header = T))
```

  * Background mutation file
  
TODO: add description about how to generate a background mutation file from driverMAPS output/ generate within diffdriver. 
```{r}
bmrf = system.file("extdata/", "example_BMRlist.Rd", package = "diffdriver")
load(bmrf)
```

  * Hotspot file
```{r}
hotf = system.file("extdata/", "example_hotspot.txt", package = "diffdriver")
head(read.table(hotf, header = T))
```

  * Directory that contains annotation files
We have two types of annotations, one has 96 annotation files: 
```{r}
annodir = "~/temp/annodir96"
list.files(annodir)
```
One has 9 annotation files:

```{r}
annodir = "~/temp/annodir9"
list.files(annodir, pattern = "_annodata.txt")
```



# Run diffdriver

```{r}
library(diffdriver)
output_dir <- "~/temp/output"
```

With signature adjustment, takes around 10min for two genes. Each additional gene will take 3-4 minutes. Please use the provide the annotation folder with 96 annotation files. `j` is the column that will be used as the phenotype, e.g j=2 means the second column in the phenotype file will be used as phenotype. 
```{r}
res <- diffdriver(genef, mutf, phenof, bmrf, j = 2, hotf, annodir = "~/temp/annodir96", k=6, BMRmode = "signature", outputdir = outputdir, outputname = "testdiffdriver_sig")
res
```

Without signature adjustment, takes around 5min for two genes. Each additional gene will take 2 minutes.Please use the provide the annotation folder with 9 annotation files. 
```{r}
res <- diffdriver(genef, mutf, phenof, bmrf, j = 2, hotf, annodir = "~/temp/annodir9", BMRmode = "regular", outputdir = outputdir, outputname = "testdiffdriver_reg")
res
```

# Plot specific genes
```{r plot}
load(file.path(outputdir,"testdiffdriver_reg_SmokingCessation_resdd.Rd"))     # load the output file from diffdriver.
g <- "PIK3CA" # only need to change gene name.
rig <- riallg[[g]]
rig$ridx <- 1:dim(rig)[1]
muts <- data.table::fread(mutf, header = T)
if (!grepl('chr', muts$Chromosome[1], fixed = T)) {muts$Chromosome <- paste0("chr",muts$Chromosome)}
muti <- na.omit(ci[rig[muts, on = c("chrom"= "Chromosome", "start" = "Position",  "ref" = "Ref",  "alt"= "Alt")], on = "SampleID"])
mutmtx <- Matrix::sparseMatrix(i = muti$ridx, j = muti$cidx, dims = c(max(rig$ridx), max(ci$cidx)))
bmrmtx= bmrallg[[g]]
ganno <- fannoallg[[g]]

diffdriver::plot_mut(mutmtx=mutmtx, canno=e, bmrmtx=bmrmtx, ganno=ganno)
```


```{r}
sessionInfo()
```
