---
title: "A tutorial to run diffdriver"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```

# Environment requirement

Need 16G memory to run diffdriver. The place that need the most memory is for parameter estimation for background mutation rate.

# Prepare input

  * Gene file
```{r}
genef = system.file("extdata", "example_gene.txt", package = "diffdriver")
gene <- read.table(genef, header = F)
head(gene)
```

  * Mutation file
```{r}
mutf = system.file("extdata/", "example_mutations.txt", package = "diffdriver")
mut <- read.table(mutf, header = T)
head(mut)
```

  * Phenotype file

The first column is sample ID with column name "SampleID", the second column is phenotype with the phenotype name as column name (note no space or in tab in column names are allowed). 
```{r}
phenof = system.file("extdata/", "example_phenotypes.txt", package = "diffdriver")
pheno <- read.table(phenof, header = T)
head(pheno)
```

  * Directory that contains annotation files

```{r}
annodir = "~/temp/annodir96"
list.files(annodir)
```


# Run diffdriver

```{r}
# library(diffdriver)
output_dir <- "~/temp/output"
```

With signature adjustment (`BMRmode = "signature"`), which is the default mode, it will need around 20min to estimate parameters in background model. Please use the provided the annotation folder with 96-annotation files when running with the "signature mode". 
```{r}
res <- diffdriver(gene = gene, mut= mut, pheno = pheno, anno_dir = "~/temp/annodir96", k=6, totalnttype = 96, BMRmode = "signature", output_dir = output_dir, output_prefix = "testdiffdriver_sig")
res
```

Without signature adjustment (`BMRmode = "regular"`), it will need around 20min to estimate parameters in background model. You can use the provided the annotation folder with 9-annotation files or 96-annotation files. The example below uses 96-annotation files. When total number of mutations is low, one should use 9-annotation files.
```{r}
res <- diffdriver(gene = gene, mut= mut, pheno = pheno,  anno_dir = "~/temp/annodir96", totalnttype = 96, BMRmode = "regular", output_dir = output_dir, output_prefix = "testdiffdriver_reg")
res
```

# Plot specific genes

To visualize the data for specific genes, diffdriver has a plotting function: 
```{r plot}
diffdriver::plot_mut(gene_name = "PIK3CA", mut= mut, pheno = pheno, totalnttype = 9, anno_dir = "~/temp/annodir9")
```

```{r}
sessionInfo()
```
